### 2018-07-10 appendix/table_correlation_nmf_lda.tex

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{llc}
\toprule
& Correlation between NMF and LDA \\
\midrule
Elections & \textbf{0.98} \\
Missing person & \textbf{0.99} \\
Former Planning minister + Former Vice-President & \textbf{0.89} \\
Current President & \textbf{0.94} \\
Social leader & \textbf{0.94} \\
Prosecutor's death & \textbf{0.83} \\
\bottomrule
\end{tabular}}
\caption{Correlation between the temporal profiles of the topics found in NMF and associated topics in LDA.}
\label{table:correlation_nmf_lda}
\end{table}