### 2018-07-10 discussion/table_gt_queries_correlation.tex

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{llccc}
\toprule
& Google Trends query & Correlation MA and Gt & MA and Twitter & Gt and Twitter \\ 
\midrule
Elections & elecciones + cambiemos + cristina kirchner + massa + randazzo & \textbf{0.81} & \textbf{0.59} & \textbf{0.75} \\
Missing person & santiago maldonado + juez otranto + patricia bullrich + gendarmería + desaparición forzada & \textbf{0.68} & \textbf{0.76} & \textbf{0.89} \\
Former Planning minister & de vido + desafuero + ministro de planificación + minnicelli + baratta & \textbf{0.92} & \textbf{0.82} & \textbf{0.87} \\
Current President & mauricio macri + cgt + reforma laboral + peña + triaca & \textbf{0.77} & \textbf{0.75} & \textbf{0.63} \\
Social leader & milagro sala + cidh + tupac amaru + pullen llermanos + morales & \textbf{0.49} & \textbf{0.25(*)} & \textbf{0.57} \\
Prosecutor's death & nisman + amia + memorándum con irán + timerman + juez bonadio & \textbf{0.56} & \textbf{0.59} & \textbf{0.75} \\
Former Vice-President & amado boudou + ciccone + ariel lijo + vandenbroele + núñez carmona & \textbf{0.90} & \textbf{0.92} & \textbf{0.97}\\
\bottomrule
\end{tabular}}

\caption{Queries performed in Google Trends in order to made up the Public Agenda. 
We also shown the correlation between the topics' temporal profiles of the Public Agenda and their counterpart in Media Agenda.
All correlation values are statistical significant ($p < 10^{-9}$), except (*) which is significant with $p < 0.05$.}
\label{table:gt_all_correlation}
\end{table}