### 2018-07-10 intro/argentinian_newspapers.tex

\subsection{Argentinian works}

% Argentinian or newspapers in spanish
\par In Argentina, the role of Mass Media has been particularly discussed during the administration of Argentina’s president \emph{Cristina Fernández de Kirchner} (from 2007 to 2015), where the government started to confront with news organizations that were critical to its management as they were an opposition party \cite{mitchelstein2017information}. Several works by E. Zunino, like \cite{zunino2010cobertura} and \cite{koziner2013cobertura}, studied the coverage given by the main newspapers of particular events which have happened during that period. 
On the other hand, Sagarzazu studied the bias present in the argentinian media respect to \emph{Hugo Chávez Frias} as a reflection of the ideology of the respective newspapers  \cite{sagarzazu2017hugo}.