### 2018-07-10 intro/our_work.tex

\subsection{Our contribution}

% Presentación y descripción de nuestro trabajo
\par In this work we propose a simple method to the study of Mass Media and audience response through topic detection algorithms.
Our work intends to contribute another quantitative approach which complements the agenda-setting theory describe above. 
\par On the other hand, we aim to take an insight about Media dynamics and Public response in order to create useful tools at the time of constructing mathematical models about the interaction between Media and Public, investigation that we started in \cite{pinto2016setting}.